### 2023-08-12 sections/3-methods.tex
updates to sieve v1 alg

Given a query $q$ and a positive integer $k$, find the $k$ closest instances in $X$.

Given a Cluster $C$, let $c$ be its center and $r$ be its radius.
Let $\delta = f(q, c)$ be the distance from the query to the cluster center.
Let $\delta_{max} = \delta + r$ be the distance from the query to the potentially farthest instance in the Cluster.
Let $\delta_{min} = max(0, \delta - r)$ be the distance from the query to the potentially closest instance in the Cluster.

@@

Create a priority queue, $candidates$, wherein the top priority candidate is the one with the smallest $\delta_{min}$.
Initialize $candidates$ with the root cluster.
Create a second priority queue, $hits$, wherein the top priority hit is the 
one with the largest $\delta$. This queue starts empty.

@@

Extract hits.

\begin{algorithm} % enter the algorithm environment
\caption{Expanding Threshold} % give the algorithm a caption
\label{alg:expanding_threshold} % and a label for \ref{} commands later in the document
\begin{algorithmic}[1] % enter the algorithmic environment
    \REQUIRE $tree$, $query$, $k$
    \STATE $candidates \leftarrow$ priority queue
    \STATE $hits \leftarrow$ priority queue
    \WHILE{$|hits| < k \lor hits.peek().\delta > candidates.peek().\delta_{min}$}
        \WHILE{$!candidates.peek().isLeaf$}
            \STATE $[l, r] \leftarrow candidates.pop().children()$
            \STATE $candidates.push([l, r])$
        \ENDWHILE
        \STATE $leaf \leftarrow candidates.pop()$
        \STATE $hits.push(leaf)$
        \WHILE{$|hits| > k$}
            \STATE $hits.pop()$
        \ENDWHILE
    \ENDWHILE
    \STATE Return $hits$
\end{algorithmic}
\end{algorithm}

\subsubsection{Sieve V1}
\label{subsubsec:methods:knn-search:sieve-v1}
Add explanation of the sieve v1 algorithm.

@@

Let $hits$ be a a fixed size priority queue with $k$ isntances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
Initialize $hits$ with the centers of $candidates$.

>>>>

The k-nearest neighbors search problem is posed as follows: given a query $q$ and a positive integer $k$, find the $k$ closest instances in $X$.

Given a Cluster $C$, let $c$ be its center and $r$ be its radius. Our $k$-nn variants make use of the following definitions:
\begin{itemize}
    \item $\delta = f(q, c)$ is the distance from the query to the cluster center.
    \item $\delta_{max} = \delta + r$ is the distance from the query to the potentially farthest instance in the Cluster.
    \item $\delta_{min} = \text{max}(0, \delta - r)$ is the distance from the query to the potentially closest instance in the Cluster.
\end{itemize}

@@

Let $candidates$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $candidates$ with the root cluster.
Let $hits$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.

@@

Extract $hits$.

\begin{algorithm} % enter the algorithm environment
\caption{Expanding Threshold} % give the algorithm a caption
\label{alg:expanding_threshold} % and a label for \ref{} commands later in the document
\begin{algorithmic}[1] % enter the algorithmic environment
    \REQUIRE $tree$, $query$, $k$
    \STATE $candidates \leftarrow$ priority queue
    \STATE $candidates.push(tree.root)$
    \STATE $hits \leftarrow$ priority queue
    \WHILE{$|hits| < k \lor hits.peek().\delta > candidates.peek().\delta_{min}$}
        \WHILE{$!candidates.peek().isLeaf$}
            \STATE $[l, r] \leftarrow candidates.pop().children()$
            \STATE $candidates.push([l, r])$
        \ENDWHILE
        \STATE $leaf \leftarrow candidates.pop()$
        \STATE $hits.push(leaf)$
        \WHILE{$|hits| > k$}
            \STATE $hits.pop()$
        \ENDWHILE
    \ENDWHILE
    \STATE Return $hits$
\end{algorithmic}
\end{algorithm}

\subsubsection{Sieve V1}
\label{subsubsec:methods:knn-search:sieve-v1}
Let $candidates$ be a vector of $clusters$s sorted by non-decreasing $\delta_{max}$. Initialize $candidates$ with  
the root cluster. 
Let $hits$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.

In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{minmax}$ less than $\tau$ 
can contain one of the $k$ nearest neighbors.

Repeat the following: 
\begin{enumerate}
\item Find the cluster $C_{\tau}$ in $candidates$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
\item Remove from $hits$ any points which are a distance greater than $\tau$ from the query. 
\item If no cluster in $candidates$ has a $\delta_{max}$ less than or equal to $\tau$, break.
\item Remove from $candidates$ any cluster whose $\delta_{min}$ is greater than $\tau$.
\item Transfer from $candidates$ to $hits$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
\item Replace all non-leaf clusters in $candidates$ with their children. 
\end{enumerate}

At this stage, if any clusters remain in $candidates$, we add all their points to $hits$. 
Extract $hits$.

\begin{algorithm} % enter the algorithm environment
    \caption{Sieve V1} % give the algorithm a caption
    \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[2] % enter the algorithmic environment
        \REQUIRE $tree$, $query$, $k$
        \STATE $candidates \leftarrow$ priority queue
        \STATE $candidates.push(tree.root)$
        \STATE $hits \leftarrow$ priority queue
        \WHILE{$|hits| < k \lor hits.peek().\delta > candidates.peek().\delta_{min}$}
            \WHILE{$!candidates.peek().isLeaf$}
                \STATE $[l, r] \leftarrow candidates.pop().children()$
                \STATE $candidates.push([l, r])$
            \ENDWHILE
            \STATE $leaf \leftarrow candidates.pop()$
            \STATE $hits.push(leaf)$
            \WHILE{$|hits| > k$}
                \STATE $hits.pop()$
            \ENDWHILE
        \ENDWHILE
        \STATE Return $hits$
    \end{algorithmic}
    \end{algorithm}

@@

Let $hits$ be a a fixed size priority queue with $k$ instances sorted by non-decreasing $f(q, x)$, i.e. distance to the query.
Initialize $hits$ with the centers of $candidates$.
